---
title: "Multiplexed CHO single cell RNA-seq"
date: "`r Sys.Date()`"
author: Clarke Lab. NIBRT
output:
  rmdformats::downcute:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: true
    highlight: tango
---

# Overview

This R notebook enables the reproduction of the analysis described in:

Tzani *et al.* Understanding the transcriptional response to ER stress in Chinese hamster ovary cells using multiplexed single cell RNA-seq (2022) [bioRxiv](bioRxivlink)

In this study, we report the use of oligonucleotide barcoding to perform multiplexed CHO cell scRNA-seq to study the impact of tunicamycin (TM), an inducer of the unfolded protein response (UPR). For this experiment, we treated a CHO-K1 GS cell line with 10μg/ml tunicamycin and acquired samples at 1, 2, 4 and 8 hr post-treatment as well as a non-treated TM- control. We transfected cells from each sample with a distinct polyadenylated ssDNA oligonucleotide barcode before pooling all samples for scRNA-seq. 

## Raw Data Availibility
The FASTQ files for this study are available from the NCBI Sequence Read Archive ID: [PRJNA821033](SRAlink)     

The associated github repository can be found [here](githublink)

# Data Analysis

## Prepare

```{r preparation, results='hide'}

#load packages
package_list <- c(
  "DropletUtils", "Matrix", "tidyverse", "readxl", "writexl", "scales", 
  "ggridges", "WGCNA", "scWGCNA", "DT", "Seurat", "viridis", "WebGestaltR", 
  "ggpubr", "patchwork", "scater","ggExtra", "SeuratWrappers", "langevitour", 
  "biomaRt","RColorBrewer", "plyr"
)
invisible(lapply(package_list, require, character.only = TRUE))


set.seed(38)

# create a directory for the output
results_dir <- "../results/"
if (!dir.exists(results_dir)) {
  dir.create(results_dir, recursive = TRUE)
}
```

## qPCR analysis of gene expression

Prior to conducting single cell analysis, qPCR was carried was to confirm that the expression of 3 genes and a spliced isoform of XBP1 were upregulated following tunicamycin addition. Here, we plot the fold changes for each treatment timepoint against the untreated control. 

```{r qpcr, fig.cap="The upregulation of Atf4, Ddit3, Hspa5 and sXbp1 in comparison to the non-TM treated control sample observed via qPCR, confirmed the induction of the UPR"}
qpcr_data <- read_excel("../data/qpcr_analysis.xlsx")

qpcr_data <- qpcr_data %>%
  pivot_longer(cols = c("Hspa5", "Ddit3", "Atf4", "sXbp1")) %>%
  mutate(fold_change = signif(value, 2))

# define color pa
col_pal_vidiris <- c(
  "#fde725", "#5ec962", "#21918c",
  "#31688e", "#440154"
)

qpcr_data %>%
  ggboxplot(
    x = "Condition", y = "fold_change",
    fill = "Condition", palette = col_pal_vidiris
  ) +
  geom_point(
    data = qpcr_data, aes(x = Condition, y = fold_change),
    position = position_jitter(width = 0),
    size = 0.2, color = "grey"
  ) +
  labs(x = "", y = "Fold change") +
  facet_wrap(~name, scales = "free_y", nrow = 2) +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text.x = element_text(face = "italic"),
    strip.background = element_blank(),
    panel.spacing = unit(1, "lines"),
    axis.text.x = element_text(angle = 35, vjust = 1, hjust = 1)
  ) +
  scale_y_continuous(breaks = pretty_breaks(n = 4, min.n = 0))

ggsave(
  filename = paste(results_dir, "Figure 1A.png", sep = ""),
  width = 4, height = 4, device = "png", dpi = 700
)
```

## Preprocess cellular RNA libraries 

In this section we preprocess the counts for the CHO cell scRNA-seq data. We first remove empty droplets and then filter cells barcodes to remove data that may arise from damaged cells.  

Load the cellular RNA data from GEM well 1 and GEM well 2. These data have already undergone filtering for empty droplets. 

```{r load_rna_data}
# GEM well 1 
bc_rank1 <- readRDS("../data/rds_files/library1_bc_rank.rds")
cell_library_1 <- readRDS("../data/rds_files/cellular_rna_lib1.rds")

# GEM well 1
bc_rank2 <- readRDS("../data/rds_files/library2_bc_rank.rds")
cell_library_2 <- readRDS("../data/rds_files/cellular_rna_lib2.rds")
```


Show the knee plots illustrating filtering of empty droplets. Function was taken from from the Kallisto | Bustools [tutorial](<https://www.kallistobus.tools/tutorials/kb_getting_started/r/kb_intro_2_r/>)
```{r knee_plot, fig.height=4, fig.width=8, fig.cap="Knee plots illustrating the removal of empty droplets for the cellular RNA data for (a) GEM well 1 and (b) GEM well 2 based on ranking the total UMI for counts per cellular barcode"}
# function to construct the plot
knee_plot <- function(bc_rank) {
  knee_plt <- tibble(
    rank = bc_rank[["rank"]],
    total = bc_rank[["total"]]
  ) %>%
    distinct() %>%
    dplyr::filter(total > 0)
  annot <- tibble(
    inflection = metadata(bc_rank)[["inflection"]],
    rank_cutoff = max(bc_rank$rank[bc_rank$total > metadata(bc_rank)[["inflection"]]])
  )
  
  p <- ggplot(knee_plt, aes(total, rank)) +
    geom_line() +
    geom_hline(aes(yintercept = rank_cutoff), data = annot, linetype = 2) +
    geom_vline(aes(xintercept = inflection), data = annot, linetype = 2) +
    scale_x_log10() +
    scale_y_log10() +
    annotation_logticks() +
    labs(y = "Rank", x = "Total UMIs")
  return(p)
}

# plot
knee_plot_lib1 <- knee_plot(bc_rank1) + 
  labs(title = "Cellular RNA library 1") + theme_bw()

knee_plot_lib2 <- knee_plot(bc_rank2) + 
  labs(title = "Cellular RNA library 2") + theme_bw()

# join and tag
knee_plot_lib1 + knee_plot_lib2 + plot_annotation(tag_levels = "a")

# save
ggsave(
  filename = paste(results_dir, "Figure S2.png", sep = ""),
  width = 8, height = 4, device = "png", dpi = 700
)
```
Next, we create a function to filter cells from both libraries based on the number of UMIs mapping to mtDNA as well as the number of genes and UMIs detected. 
```{r rna_filter_function}
filter_cells <- function(object) {
  
  # run miQC
  object <- RunMiQC(object,
    percent.mt = "percent.mt",
    nFeature_RNA = "nFeature_RNA",
    posterior.cutoff = 0.9,
    model.slot = "flexmix_model"
  )
  
  # capture the plot
  mtdna_plot <- PlotMiQC(object, color.by = "miQC.keep")
  
  # remove poor quality cells
  object <- subset(object, subset = miQC.keep == "keep")
  
  # modify miQC output plot
  mtdna_plot$layers[c(1, 2, 3)] <- NULL
  mtdna_plot <- mtdna_plot + theme_bw() +
    scale_color_manual(values = c("#888888", "#117733")) +
    # scale_color_brewer(palette = "Dark2") +
    geom_point(size = 0.2, alpha = 0.5) +
    labs(x = "# Genes", y = "% mtDNA UMIs", color = "") +
    guides(colour = guide_legend(override.aes = list(size = 5))) +
    annotate("text", -Inf, Inf,
      hjust = -1.70, vjust = 1.5, color = "#117733",
      label = paste0(dim(object)[2], " cells retained")
    ) +
    theme(
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      legend.position = "bottom"
    )

  # thresholds for nUMI
  log_umi_counts <- log10(object@meta.data$nCount_RNA)
  low_count_threshold <- median(log_umi_counts) - (3 * mad(log_umi_counts))
  high_count_threshold <- median(log_umi_counts) + (3 * mad(log_umi_counts))

  # thresholds for detected genes
  log_detect_genes <- log10(object@meta.data$nFeature_RNA)
  low_feature_threshold <- median(log_detect_genes) - (3 * mad(log_detect_genes))
  high_feature_threshold <- median(log_detect_genes) + (3 * mad(log_detect_genes))

  # plot data for UMI and detected genes results
  umi_v_feature_plot_data <- object@meta.data %>%
    mutate(miQC.keep = case_when(
      nFeature_RNA > 10^low_feature_threshold &
        nFeature_RNA < 10^high_feature_threshold &
        nCount_RNA > 10^low_count_threshold &
        nCount_RNA < 10^high_count_threshold ~ "keep",
      TRUE ~ "discard"
    ))
  
  # filter cells
  object <- subset(
    object,
    nFeature_RNA > 10^low_feature_threshold &
      nFeature_RNA < 10^high_feature_threshold &
      nCount_RNA > 10^low_count_threshold &
      nCount_RNA < 10^high_count_threshold
  )

  # plot UMI and detected genes results
  umi_v_feature_plot <- umi_v_feature_plot_data %>%
    ggplot(aes(x = log10(nCount_RNA), y = log10(nFeature_RNA), color = miQC.keep)) +
    geom_point(size = 0.2, alpha = 0.5) +
    scale_color_manual(values = c("#888888", "#117733")) +
    # geom_smooth(method = "lm") +
    geom_hline(aes(yintercept = low_feature_threshold), colour = "grey", linetype = 2) +
    geom_hline(aes(yintercept = high_feature_threshold), colour = "grey", linetype = 2) +
    geom_vline(aes(xintercept = high_count_threshold), colour = "grey", linetype = 2) +
    geom_vline(aes(xintercept = low_count_threshold), colour = "grey", linetype = 2) +
    theme_bw() +
    labs(x = "Log10 # UMIs", y = "Log10 # Genes", color = "") +
    guides(colour = guide_legend(override.aes = list(size = 5))) +
    annotate("text", -Inf, Inf,
      hjust = -0.1, vjust = 1.5, color = "#117733",
      label = paste0(dim(object)[2], " cells retained")
    ) +
    theme(
      panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
      legend.position = "bottom"
    )


  # show the filtered cell summary
  print(object)

  # results list
  result_list <- list(
    "mtdna_plot" = mtdna_plot,
    "filtered_object" = object,
    "umi_feature_plot" = umi_v_feature_plot,
    "umi_feature_plot_data" = umi_v_feature_plot_data
  )

  return(result_list)
}
```

Specify the mitochondrial protein coding genes from the CHO-K1 genome
```{r mito_gene_list}
mito <- c("ENSCGRG00000000006.1","ENSCGRG00000000010.1","ENSCGRG00000000016.1",
          "ENSCGRG00000000019.1","ENSCGRG00000000021.1","ENSCGRG00000000022.1",
          "ENSCGRG00000000023.1","ENSCGRG00000000025.1","ENSCGRG00000000027.1",
          "ENSCGRG00000000028.1","ENSCGRG00000000032.1","ENSCGRG00000000033.1",
          "ENSCGRG00000000035.1")
```

Preprocess GEM well 1 RNA data.  

```{r preprocess_rna_1, class.source="watch-out"}
# determine the percentage mapping to mtDNA
cell_library_1[["percent.mt"]] <- PercentageFeatureSet(cell_library_1, 
                                                       features = mito)
# preprocess
cell_rna_library_1_filter_output <- filter_cells(cell_library_1)

cell_library_1 <- cell_rna_library_1_filter_output$filtered_object
```

Preprocess GEM well 2 RNA data.. 

```{r preprocess_rna_2, warning=FALSE}
# determine the percentage mapping to mtDNA
cell_library_2[["percent.mt"]] <- PercentageFeatureSet(cell_library_2, 
                                                       features = mito)
# preprocess
cell_rna_library_2_filter_output <- filter_cells(cell_library_2)

cell_library_2 <- cell_rna_library_2_filter_output$filtered_object
```

Plot the results of cellular RNA preprocessing for GEM well 1 and 2   

```{r plot_preprocessing_output, fig.height=8, fig.width=10, fig.cap="miQC was used to assess the relationship between the number of UMIs originating from mtDNA and the number genes detected to identify low quality cells in the cellular RNA libraries from (a) GEM well 1 and (c) GEM well 2. If the number of UMIs or number of genes detected per cell was beyond 3 MAD of the median of the population of (b) GEM well 1 or (d) GEM well 2 those cells were also discarded."}
(
  (cell_rna_library_1_filter_output$mtdna_plot + plot_spacer() + 
    cell_rna_library_1_filter_output$umi_feature_plot + 
    plot_layout(widths = c(2,0.1,2))
   ) 
  / 
(
  cell_rna_library_2_filter_output$mtdna_plot + plot_spacer()  + 
   cell_rna_library_2_filter_output$umi_feature_plot + 
   plot_layout(widths = c(2,0.1,2))
  )
) + 
  plot_annotation(tag_levels = "a") + plot_layout(guides = "collect") & theme(legend.position = "bottom")

ggsave(filename = paste(results_dir, "Figure S3.png", sep = ""), 
       width = 9, height = 9, device = "png", dpi = 700)
```

## Sample classification using SBO data

First, we define a function to integrate the SBO with its cellular RNA data.  


```{r sbo_integration_fuction}
sbo_integration <- function(object, sbo_matrix, library) {

  # extract the RNA counts from the Seurat object
  seurat_count_matrix <- as.matrix(GetAssayData(
    object = object,
    slot = "counts"
  ))

  # determine the cell barcodes shared between SBO and RNA
  joint_barcodes <- intersect(
    colnames(sbo_matrix),
    colnames(seurat_count_matrix)
  )

  print(paste(length(joint_barcodes), "barcodes matched between mRNA and SBO library"))

  # create a RNA count matrix containing only the shared barcodes
  matched_count_matrix <- seurat_count_matrix[, joint_barcodes]

  # create a new Seurat object for the RNA counts
  new_object <- CreateSeuratObject(
    counts = matched_count_matrix,
    project = library
  )

  # set the percent mt for new object using the previous values
  new_object[["percent.mt"]] <- object$percent.mt[joint_barcodes]

  # add meta data for the library
  new_object@meta.data$library <- rep(library, dim(new_object)[2])

  # add a new assay to hold the SBO counts
  sbo_matrix <- as.matrix(sbo_matrix[, joint_barcodes]) # intersect

  new_object[["SBO"]] <- CreateAssayObject(counts = sbo_matrix)

  return(new_object)
}
```

Integrate RNA and SBO counts for GEM well 1
```{r integrate_sbo_1, warning=FALSE, message=FALSE}
# load the SBO counts for GEM well 2
sbo_lib_1_data <- readRDS("../data/rds_files/sbo_lib_1.data.rds")

# integrate
cho_cell_sbo_lib_1 <- sbo_integration(cell_library_1, sbo_lib_1_data, "library 1")

#summary
cho_cell_sbo_lib_1

# normalise SBO counts 
cho_cell_sbo_lib_1 <- NormalizeData(cho_cell_sbo_lib_1, assay = "SBO", 
                               normalization.method = "CLR")
```

```{r ntegrate_sbo_2, message=FALSE, warning=FALSE}
# load the SBO counts for GEM well 2
sbo_lib_2_data <- readRDS("../data/rds_files/sbo_lib_2.data.rds")

# integrate
cho_cell_sbo_lib_2 <- sbo_integration(cell_library_2, sbo_lib_2_data, "library 2")

#summary
cho_cell_sbo_lib_2

# normalise SBO counts 
cho_cell_sbo_lib_2<- NormalizeData(cho_cell_sbo_lib_2, assay = "SBO", 
                               normalization.method = "CLR")
```

Next, we need to determine a postive quartile threshold for SBO classification for the 2 GEM well libraries.  

```{r evaluate_positive_quantile}
eval_pq <- function(object, library_id) {
  
  eval_pq_out <- tibble()

  for (i in seq(from = 0.9, to = 0.999999999, by = 0.005)) {
    object <- HTODemux(object,
      assay = "SBO",
      positive.quantile = i,
      verbose = F,
      kfunc = "kmeans", init = 6
    )

    sbo_classification <- as.data.frame.matrix(table(
      object$SBO_maxID,
      object$SBO_classification.global
    )) %>%
      summarise(max.singlet = sum(Singlet)) %>%
      mutate(threshold = i)

    eval_pq_out <- bind_rows(eval_pq_out, sbo_classification) %>%
      mutate(library = library_id)
  }

  eval_pq_plot <- eval_pq_out %>%
    ggplot(aes(x = threshold, y = cells, fill = SBO_classification.global)) +
    geom_bar(position = "fill", stat = "identity") +
    theme_bw() +
    scale_fill_viridis(discrete = T, direction = -1, option = "plasma") +
    labs(y = "# Cells", x = "HTODemux positive quantile treshold", fill = "SBO identification", title = library_id)

  result_list <- list(
    "eval_pq_out" = eval_pq_out,
    "eval_pq_plot" = eval_pq_plot
  )

  return(result_list)
}
```

Here we select a balenced cutoff for both GEM wells to optimise the total number of singlet cells identified.   

```{r evaluate_sbo_threshold, fig.height=5, fig.width=7, fig.cap="A range of positive quantile value from 0.9 to 0.999999 were assessed to identify a value for both GEM well 1 and GEM well 2 datasets. 0.94 was selected as the classification threshold to maximise the number of singlet cells."}
eval_pq_lib1 <- eval_pq(cho_cell_sbo_lib_1, "GEM well 1")

eval_pq_lib2 <- eval_pq(cho_cell_sbo_lib_2, "GEM well 2")

bind_rows(eval_pq_lib1$eval_pq_out, eval_pq_lib2$eval_pq_out) %>%
  ggplot(aes(x = threshold, y = max.singlet, fill = library, color = library)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept = 0.94, linetype = 11, size = 0.5, color = "grey") +
  geom_label(aes(x = 0.94, label = "HTODemux\nPostive Quantile\n= 0.94", y = 2550), inherit.aes = F, colour = "dark grey") +
  scale_color_brewer(palette = "Dark2", direction = -1) +
  labs(x = "Postive Quantile", y = "# Singlet cells", fill = "", color = "") +
  theme_bw()

ggsave(filename = paste(results_dir, "Figure S4.png", sep = ""), width = 6, height = 4, device = "png", dpi = 700)
```

Classify GEM well 1 samples.  

```{r classify_rna_lib_1}
# classify
cho_cell_sbo_lib_1 <- HTODemux(cho_cell_sbo_lib_1,
  assay = "SBO",
  positive.quantile = 0.94, kfunc = "kmeans", init = 6
)

lib1_sbo_classification <- as.data.frame.matrix(
  table(
    cho_cell_sbo_lib_1$SBO_maxID,
    cho_cell_sbo_lib_1$SBO_classification.global
  )
) %>%
  as_tibble(rownames = "Sample") %>%
  mutate(`Sample label` = case_when(
    Sample == "control-CCGGACTT" ~ "TM-",
    Sample == "1-hours-treated-GCTTGCCT" ~ "TM+ 1hr",
    Sample == "2-hours-treated-GAACTCCC" ~ "TM+ 2hr",
    Sample == "4-hours-treated-GTCAGGAT" ~ "TM+ 4hr",
    Sample == "8-hours-treated-GTATATCC" ~ "TM+ 8hr"
  )) %>%
  dplyr::rename(`Sample ID` = Sample) %>%
  dplyr::select(
    `Sample ID`, `Sample label`,
    `Doublet`, `Negative`, `Singlet`
  ) %>%
  arrange(`Sample label`)

# output table
lib1_sbo_classification %>%
  DT::datatable()

# capture data for merged classification plot
classification_library_1 <- data.frame(table(
  cho_cell_sbo_lib_1$SBO_maxID,
  cho_cell_sbo_lib_1$SBO_classification.global
)) %>%
  mutate(library = "GEM well 1")

# remove cells with no classified SBO
rna_sbo_lib_1_subset <- subset(cho_cell_sbo_lib_1,
  idents = "Negative", invert = T
)
```

Classify GEM well 1 samples. 

```{r classify_rna_lib_2}
# classify
cho_cell_sbo_lib_2 <- HTODemux(cho_cell_sbo_lib_2,
  assay = "SBO",
  positive.quantile = 0.94, kfunc = "kmeans", init = 6
)


lib2_sbo_classification <- as.data.frame.matrix(
  table(
    cho_cell_sbo_lib_2$SBO_maxID,
    cho_cell_sbo_lib_2$SBO_classification.global
  )
) %>%
  as_tibble(rownames = "Sample") %>%
  mutate(`Sample label` = case_when(
    Sample == "control-CCGGACTT" ~ "TM-",
    Sample == "1-hours-treated-GCTTGCCT" ~ "TM+ 1hr",
    Sample == "2-hours-treated-GAACTCCC" ~ "TM+ 2hr",
    Sample == "4-hours-treated-GTCAGGAT" ~ "TM+ 4hr",
    Sample == "8-hours-treated-GTATATCC" ~ "TM+ 8hr"
  )) %>%
  dplyr::rename(`Sample ID` = Sample) %>%
  dplyr::select(
    `Sample ID`, `Sample label`,
    `Doublet`, `Negative`, `Singlet`
  ) %>%
  arrange(`Sample label`)

lib2_sbo_classification %>%
  DT::datatable()

# capture data for merged classification plot
classification_library_2 <- data.frame(
  table(
    cho_cell_sbo_lib_2$SBO_maxID,
    cho_cell_sbo_lib_2$SBO_classification.global
  )
) %>%
  mutate(library = "GEM well 2")

# remove cells with no classified SBO
rna_sbo_lib_2_subset <- subset(cho_cell_sbo_lib_2,
  idents = "Negative", invert = T
)
```

Summarise SBO classifications and write to file.  


```{r classification_summary}
combined_sbo_classification <- tibble(
  `Sample ID` = lib1_sbo_classification$`Sample ID`,
  `Sample label` = lib1_sbo_classification$`Sample label`,
  Doublet = lib1_sbo_classification$Doublet + lib2_sbo_classification$Doublet,
  Negative = lib1_sbo_classification$Negative + lib2_sbo_classification$Negative,
  Singlet = lib1_sbo_classification$Singlet + lib2_sbo_classification$Singlet
)

combined_sbo_classification %>%
  DT::datatable()

# write supplementary table
filename <- paste(results_dir, "Table S3.xlsx",
  sep = ""
)

write_xlsx(list(
  a = combined_sbo_classification,
  b = lib1_sbo_classification,
  c = lib2_sbo_classification
),
path = filename,
format_headers = TRUE
)
```

Combine the two GEM well libraries.  

```{r combine_tagged_cells, message=FALSE, warning=FALSE}
# merge non-negative cells
rna_sbo_combined <- merge(rna_sbo_lib_1_subset,
  y = c(rna_sbo_lib_2_subset),
  add.cell.ids = c("lib1", "lib2"),
  project = "TM_analysis", merge.data = TRUE
)

# summary
rna_sbo_combined
```

Visualise the SBO counts using t-Stochastic neighbour embedding.  

```{r plot_sbo_tsne, message=FALSE, warning=FALSE, fig.cap="t-Stochastic neighbour embedding (tSNE) dimensionality reduction and visualisation of the combined GEM well 1 and GEM well 2 UMI-SBO count matrices. HTODemux classifications are shown along with those cells identified as doublets.t-Stochastic neighbour embedding (tSNE) dimensionality reduction and visualisation of the combined GEM well 1 and GEM well 2 UMI-SBO count matrices. HTODemux classifications are shown along with those cells identified as doublets."}

sbo_dist_mtx <- as.matrix(dist(t(GetAssayData(object = rna_sbo_combined, assay = "SBO"))))

rna_sbo_combined <- RunTSNE(rna_sbo_combined, distance.matrix = sbo_dist_mtx, 
                            perplexity = 100)

sbo_combined_tsne_plot <- DimPlot(rna_sbo_combined)

figure_2b <- sbo_combined_tsne_plot$data %>%
  mutate(tag = case_when(
    ident == "control-CCGGACTT" ~ "TM- [CCGGACTT]",
    ident == "1-hours-treated-GCTTGCCT" ~ "TM+ 1hr [GCTTGCCT]",
    ident == "2-hours-treated-GAACTCCC" ~ "TM+ 2hr [GAACTCCCT]",
    ident == "4-hours-treated-GTCAGGAT" ~ "TM+ 4hr [GTCAGGAT]",
    ident == "8-hours-treated-GTATATCC" ~ "TM+ 8hr [GTATATCC]",
    ident == "Doublet" ~ "Doublet",
  )) %>%
  ggplot(aes(x = tSNE_1, y = tSNE_2, color = tag)) +
  geom_point(size = 0.1) +
  theme_bw() +
  scale_color_manual(values = c("red", "#fde725", "#5ec962", "#21918c", "#31688e", "#440154")) +
  labs(x = "tSNE 1", y = "tSNE 2", color = "Sample [barcode]") +
  guides(color = guide_legend(override.aes = list(size = 2)))

figure_2b
ggsave(plot = figure_2b, filename = paste(results_dir, "Figure_S5.png", sep = ""), width = 6, height = 4, device = "png", dpi = 700)
```

Show overall sample classifications.  

```{r plot_classification_supp, fig.height=5, fig.width=10, warning=TRUE, fig.cap="HTODemux with a positive quantile threshold = 0.94 was used to identify a total of (a) 2,645 and 2,669 singlets from GEM well 1 and GEM well 2 respectively. The (b,c) sample classifications across the 2 GEM wells were comparable. The number of singlet classifications are shown in white; the number of negative and doublet classifications are provided in Table S3. For the TM+ 4hr sample, a comparable number of SBO negative cells was observed for GEM well 1 (n = 141) and GEM well 2 (n = 190). While the origin of this variation is unknown these data indicate that the difference in detection rate of TM+ 4hr arose from the experimental steps conducted prior to cell isolation."}
overall_classification_counts <- bind_rows(classification_library_1, classification_library_2) %>%
  dplyr::rename("SBO_classification.global" = Var2) %>%
   mutate(gem_well = case_when(library == "library 1" ~ "GEM well 1", 
                              TRUE ~ "GEM well 2")) %>%
  mutate(ident = case_when(
    Var1 == "control-CCGGACTT" ~ "TM- [CCGGACTT]",
    Var1 == "1-hours-treated-GCTTGCCT" ~ "TM+ 1hr [GCTTGCCT]",
    Var1 == "2-hours-treated-GAACTCCC" ~ "TM+ 2hr [GAACTCCCT]",
    Var1 == "4-hours-treated-GTCAGGAT" ~ "TM+ 4hr [GTCAGGAT]",
    Var1 == "8-hours-treated-GTATATCC" ~ "TM+ 8hr [GTATATCC]",
    Var1 == "Doublet" ~ "Doublet",
  )) %>%
dplyr::group_by(library, SBO_classification.global) %>%
  dplyr::summarise(total_cells = sum(Freq))

fig_s6_a <- overall_classification_counts  %>%
  ggplot(aes(x = library, y = total_cells, fill = SBO_classification.global)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(aes(x= library, y=3000, fill=SBO_classification.global, label=total_cells), 
            data = overall_classification_counts  %>% filter(SBO_classification.global =="Singlet"), 
            size = 3, position = position_stack(vjust = 0.5),
            color="white") + 
  labs(y = "# Cells", x = "", fill = "") +
  theme_bw() +
  theme(
    legend.position = "right",
    axis.text.x = element_text(angle = 35, vjust = 1, hjust = 1)
  ) +
  scale_fill_viridis(discrete = T, direction = -1, option = "plasma")



gem_well_classification_counts <-bind_rows(classification_library_1, classification_library_2) %>%
  dplyr::rename("SBO_classification.global" = Var2) %>%
  mutate(ident = case_when(
    Var1 == "control-CCGGACTT" ~ "TM- [CCGGACTT]",
    Var1 == "1-hours-treated-GCTTGCCT" ~ "TM+ 1hr [GCTTGCCT]",
    Var1 == "2-hours-treated-GAACTCCC" ~ "TM+ 2hr [GAACTCCCT]",
    Var1 == "4-hours-treated-GTCAGGAT" ~ "TM+ 4hr [GTCAGGAT]",
    Var1 == "8-hours-treated-GTATATCC" ~ "TM+ 8hr [GTATATCC]",
    Var1 == "Doublet" ~ "Doublet",
  ))
  

fig_s6_b <- gem_well_classification_counts  %>%
  ggplot(aes(x=ident, y=Freq, fill=SBO_classification.global, label=Freq)) +
  geom_bar(stat="identity",position = "stack", width = 0.75) +
  geom_text(aes(x=ident, y=500, fill=SBO_classification.global, label=Freq), 
            data = gem_well_classification_counts %>% filter(SBO_classification.global =="Singlet"), 
            size = 3, position = position_stack(vjust = 0.5),
            color="white") +
  labs(x="", y="# Cells", fill="") +
  facet_wrap(~library, nrow = 2) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 35, vjust = 1, hjust = 1), 
        strip.background = element_blank()) +
  scale_fill_viridis(discrete = T, direction = -1, option = "plasma")


fig_s6_a + fig_s6_b + plot_annotation(tag_levels = "a") + 
  plot_layout(guides = "collect", widths = c(0.4,1)) & 
  theme(legend.position = "bottom")

ggsave(filename = paste(results_dir, "Figure S6.png", sep = ""), 
       width = 6, height = 5, device = "png", dpi = 700)
```


```{r combined_classification_count_plot, fig.cap="5,314 cells were confidently assigned to one of the 5 samples."}
gem_lib_singlet_counts <- bind_rows(classification_library_1, classification_library_2) %>%
  mutate(ident = case_when(
    Var1 == "control-CCGGACTT" ~ "TM-",
    Var1 == "1-hours-treated-GCTTGCCT" ~ "TM+ 1hr",
    Var1 == "2-hours-treated-GAACTCCC" ~ "TM+ 2hr",
    Var1 == "4-hours-treated-GTCAGGAT" ~ "TM+ 4hr",
    Var1 == "8-hours-treated-GTATATCC" ~ "TM+ 8hr",
    Var1 == "Doublet" ~ "Doublet",
  )) %>%
  dplyr::group_by(ident, Var2) %>%
  dplyr::summarise(total_cells = sum(Freq))


gem_lib_singlet_counts %>%
  ggplot(aes(x=ident, y=total_cells, fill=Var2)) +
  geom_bar(stat="identity",position = "stack", width = 0.75) +
  geom_text(aes(x=ident, y=1000, label=total_cells), 
            data = gem_lib_singlet_counts %>% 
              filter(Var2 =="Singlet"), 
            size = 3, 
            position = position_stack(vjust = 0.5),
            color="white") +
  labs(x="", y="# Cells", fill="", title="SBO sample classification") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 35, vjust = 1, hjust = 1), legend.position = "bottom") +
  scale_fill_viridis(discrete = T, direction = -1, option = "plasma")

ggsave(filename = paste(results_dir, "Figure S4.png", sep = ""), width = 4, height = 3, device = "png", dpi = 700)
```

Select cells with only singlet identifications and merge the data

```{r merge_gem_wells, message=FALSE, warning=FALSE}
# select only singlet cells from both libraries
Idents(cho_cell_sbo_lib_1) <- "SBO_classification.global"
cho_1.singlet <- subset(cho_cell_sbo_lib_1, idents = "Singlet")
  
Idents(cho_cell_sbo_lib_2) <- "SBO_classification.global"
cho_2.singlet <- subset(cho_cell_sbo_lib_2, idents = "Singlet")

# merge
classified_cells <- merge(cho_1.singlet,
  y = c(cho_2.singlet),
  add.cell.ids = c("lib1", "lib2"),
  project = "TM_analysis", merge.data = TRUE
)

# total cells
paste0(dim(classified_cells)[2], " cells with  sample classifications")
```

## Dimensionality reduction and visualisation

```{r cell_cycle_classification, message=FALSE, warning=FALSE, include=FALSE}
# run SCT and regress the effect of variation from the number of genes detected
classified_cells <- SCTransform(
  classified_cells,
  assay = "RNA",
  new.assay.name = "SCT",
  vars.to.regress = c("nFeature_RNA"), 
  variable.features.n = 3000
)

# load cell cycle gene annotations
cgr_s_phase <- read.csv(file = "../data/s_ensembl_genelist.csv")$Gene.stable.ID
cgr_g2m_phase <- read.csv(file = "../data/g2m_ensembl_genelist.csv")$Gene.stable.ID

# classify CC phase 
classified_cells <- CellCycleScoring(
  classified_cells,
  assay = "SCT",
  s.features = cgr_s_phase,
  g2m.features = cgr_g2m_phase,
  set.ident = TRUE
)
```


```{r sctransform, message=FALSE, warning=FALSE, include=FALSE}
# regress the effect of CC and number of genes detected
# note this is carried out on the RNA assay
classified_cells <- SCTransform(
  classified_cells,
  assay = "RNA",
  new.assay.name = "SCT",
  vars.to.regress = c("S.Score", "G2M.Score","nFeature_RNA"), variable.features.n = 3000,
  verbose = TRUE
)

# modify sample names
sample_information <- classified_cells@meta.data %>%
  mutate(treatment = case_when(
    SBO_maxID == "control-CCGGACTT" ~ "TM-",
    SBO_maxID == "1-hours-treated-GCTTGCCT" ~ "TM+ 1hr",
    SBO_maxID == "2-hours-treated-GAACTCCC" ~ "TM+ 2hr",
    SBO_maxID == "4-hours-treated-GTCAGGAT" ~ "TM+ 4hr",
    SBO_maxID == "8-hours-treated-GTATATCC" ~ "TM+ 8hr",
  ))

classified_cells <- AddMetaData(classified_cells, metadata = sample_information$treatment, "treatment")
```

```{r fig.cap="The Seurat CellCycleScoring function was used to classify the phase of cell cycle for the 5,314 cells based on the expression of genes associated with the S phase and the G2M phase (a cell not classified in either phase is designated as G1). The proportion of cells predicted to be in each phase for each of the 5 samples is shown in white. The likelihood of each cell being either S or G2M is also outputted by the algorithm and was used to regress the effect of cell cycle from the data during SCTransform normalisation."}

classified_cells@meta.data %>%
  group_by(treatment, Phase) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::mutate(freq = n / sum(n)) %>%
  ggplot(aes(x = treatment, y = freq, fill = Phase)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_brewer(palette = "Dark2",direction = 1) +
  geom_text(aes(label=paste0(sprintf("%1.1f", freq*100),"%")),
                     position=position_stack(vjust=0.5), colour="white", size = 3) +
  #scale_fill_manual(values = c("#CC6677", "#44AA99", "#88CCEE")) +
  theme_bw() +
  labs(x="",y = "Proportion of cells", fill = "Phase")

ggsave(filename = paste(results_dir, "Figure S7.png", sep = ""), width = 5, height = 3, device = "png", dpi = 700)
```

Run principal components analysis

```{r pca, fig.cap="Principal components analysis was performed on the SCTransform normalised gene expression data (cells = 5,314, genes = 3000). The first 20 principal components were retained for the UMAP visualisation."}
classified_cells <- RunPCA(object = classified_cells, verbose = FALSE, npcs = 50,seed.use = 42)

figure_s5 <- ElbowPlot(classified_cells,ndims = 50)

figure_s5$data %>%
  ggplot(aes(x=dims, y=stdev)) +
  geom_line() + 
  geom_point() +
  geom_vline(xintercept = 20, linetype=11, size=0.2, color="red") +
  theme_bw() +
  labs(x="Principal components", y="Standard Deviation")

ggsave(filename = paste(results_dir, "Figure S5.png", sep = ""), width = 5, height = 5, device = "png", dpi = 700)
```

Run UMAP

```{r run_umap, fig.height=8, fig.width=5}
classified_cells <- RunUMAP(object = classified_cells, 
                            dims = 1:20)
```


```{r plot_lib_cc, fig.height=3, fig.width=5}
### Supplementary Figure 6
# overlay library
seurat_dim_plot_lib <- DimPlot(classified_cells, reduction = "umap", group.by = c("library"), pt.size = 0.5)
# overlay cell cycle phase
seurat_dim_plot_cc <- DimPlot(classified_cells, reduction = "umap", group.by = c("Phase"), pt.size = 0.5)

cbp1 <- c(
  "#E69F00", "#009E73",
  "#F0E442", "#0072B2", "#D55E00", "#CC79A7"
)

figure_s6a <- seurat_dim_plot_lib$data %>%
   mutate(gem_well = case_when(library == "library 1" ~ "GEM well 1", 
                              TRUE ~ "GEM well 2")) %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = gem_well)) +
  geom_point(size = 0.2) +
  theme_bw() +
  scale_color_manual(values = cbp1) +
  labs(x = "UMAP 1", y = "UMAP 2", color = "") +
  guides(color = guide_legend(override.aes = list(size = 2)))

figure_s6b <- seurat_dim_plot_cc$data %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = Phase)) +
  geom_point(size = 0.2) +
  theme_bw() +
  scale_colour_brewer(palette = "Dark2",direction = 1) +
  # scale_color_manual(values = cbp1) +
  labs(x = "UMAP 1", y = "UMAP 2", color = "") +
  guides(color = guide_legend(override.aes = list(size = 2)))


(figure_s6a + plot_spacer() + figure_s6b) + plot_layout(widths =  c(4,0.2,4)) + plot_annotation(tag_levels = 'a') & theme(legend.position = "bottom")

ggsave(filename = paste(results_dir, "Figure S6.png", sep = ""), 
       width = 7, height = 4, device = "png", dpi = 700)
```

```{r fig.height=4, fig.width=5}
sbo_tagged_cells_umap_plot <- DimPlot(classified_cells, group.by = c("treatment"), combine = FALSE)

viridis_cp <- c("#fde725", "#5ec962", "#21918c", "#31688e", "#440154")

tm_umap <- sbo_tagged_cells_umap_plot[[1]]$data %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = treatment)) +
  geom_point(size = 0.5) +
  theme_bw() +
  coord_fixed() +
  #scale_color_brewer(palette = "Set1") +
  scale_color_manual(values = viridis_cp) +
  labs(x = "UMAP 1", y = "UMAP 2", color = "Treatment") +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  theme(
    plot.title = element_text(size = 10, face = "bold.italic"),
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )
tm_umap

ggsave(plot= tm_umap, filename = paste(results_dir, "Figure 2d.png", sep = ""), width = 5, height = 4, device = "png", dpi = 700)
```

```{r fig.height=3, fig.width=3}
classified_cells.3d <- RunUMAP(object = classified_cells, 
                            dims = 1:30, n.components = 3)

langevitour(as.data.frame(classified_cells.3d[["umap"]]@cell.embeddings), 
            classified_cells$treatment, pointSize = 2)
```

Plot ER stress marker genes

```{r er_marker_umap}
hspa5_umap <- FeaturePlot(classified_cells, features = c("ENSCGRG00015003428.1"))$data %>%
  mutate(gene = "Hspa5") %>%
  dplyr::rename(expression = "ENSCGRG00015003428.1") %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = expression)) +
  geom_point(size = 0.01) +
  scale_color_viridis(option = "turbo", direction = 1) +
  theme_bw() +
  coord_fixed() +
  # scale_x_reverse() +
  labs(x = "UMAP 2", y = "UMAP 1", color = "Hspa5 \nExpression \nlevel") +
  theme(
    plot.title = element_text(size = 10, face = "bold.italic"),
    legend.position = "right",
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 7)
  )

ddit3_umap <- FeaturePlot(classified_cells, features = c("ENSCGRG00015028401.1"))$data %>%
  mutate(gene = "Ddit3") %>%
  dplyr::rename(expression = "ENSCGRG00015028401.1") %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = expression)) +
  geom_point(size = 0.1) +
  scale_color_viridis(option = "turbo") +
  theme_bw() +
  coord_fixed() +
  # scale_x_reverse() +
  labs(x = "UMAP 2", y = "UMAP 1", color = "Ddit3 \nExpression \nlevel") +
  theme(
    plot.title = element_text(size = 10, face = "bold.italic"),
    legend.position = "right",
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 7)
  )

atf4_umap <- FeaturePlot(classified_cells, features = c("ENSCGRG00015004082.1"))$data %>%
  mutate(gene = "Atf4") %>%
  dplyr::rename(expression = "ENSCGRG00015004082.1") %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = expression)) +
  geom_point(size = 0.1) +
  scale_color_viridis(option = "turbo") +
  theme_bw() +
  coord_fixed() +
  # scale_x_reverse() +
  labs(x = "UMAP 2", y = "UMAP 1", color = "Atf4 \nExpression \nlevel") +
  theme(
    plot.title = element_text(size = 10, face = "bold.italic"),
    legend.position = "right",
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 7)
  )
```

```{r er_marker_violin, fig.width=12}
standard_normalisation <- NormalizeData(classified_cells,
  normalization.method = "LogNormalize",
  scale.factor = 10000, assay = "RNA"
)

hspa5_violin <- VlnPlot(standard_normalisation,
  features = c("ENSCGRG00015003428.1"),
  group.by = "treatment",
  assay = "RNA"
)$data %>%
  mutate(gene = "Hspa5") %>%
  dplyr::rename(expression = "ENSCGRG00015003428.1") %>%
  ggplot(aes(x = ident, y = `expression`, fill = ident)) +
  geom_violin() +
  geom_jitter(size = 0.1, alpha = 0.1, color = "black") +
  scale_fill_manual(values = c("#fde725", "#5ec962", 
                               "#21918c", "#31688e", "#440154")) +
  theme_bw() +
  labs(x = "", y = "Expression Level", subtitle = "Hspa5") +
  theme(
    plot.subtitle = element_text(face = "italic"),
    legend.position = "none",
    strip.text.x = element_text(face = "bold.italic"),
    strip.background = element_blank(),
    panel.spacing = unit(2, "lines"),
    axis.text.x = element_text(angle = 35, vjust = 1, hjust = 1)
  ) +
  scale_y_continuous(breaks = pretty_breaks(n = 4, min.n = 0))

ddit3_violin <- VlnPlot(standard_normalisation, 
                        features = c("ENSCGRG00015028401.1"), 
                        group.by = "treatment", assay = "RNA")$data %>%
  mutate(gene = "Ddit3") %>%
  dplyr::rename(expression = "ENSCGRG00015028401.1") %>%
  ggplot(aes(x = ident, y = `expression`, fill = ident)) +
  geom_violin() +
  geom_jitter(size = 0.1, alpha = 0.1) +
  scale_fill_manual(values = c("#fde725", "#5ec962", 
                               "#21918c", "#31688e", "#440154")) +
  theme_bw() +
  labs(x = "", y = "Expression Level", subtitle = "Ddit3") +
  theme(
    plot.subtitle = element_text(face = "italic"),
    legend.position = "none",
    strip.text.x = element_text(face = "bold.italic"),
    strip.background = element_blank(),
    axis.text.x = element_text(angle = 35, vjust = 1, hjust = 1)
  ) +
  scale_y_continuous(breaks = pretty_breaks(n = 4, min.n = 0))

atf4_violin <- VlnPlot(standard_normalisation, 
                       features = c("ENSCGRG00015004082.1"), 
                       group.by = "treatment", assay = "RNA")$data %>%
  mutate(gene = "Atf4") %>%
  dplyr::rename(expression = "ENSCGRG00015004082.1") %>%
  ggplot(aes(x = ident, y = `expression`, fill = ident)) +
  geom_violin() +
  geom_jitter(size = 0.1, alpha = 0.1) +
  scale_fill_manual(values = c("#fde725", "#5ec962", 
                               "#21918c", "#31688e", "#440154")) +
  theme_bw() +
  labs(x = "", y = "Expression Level", subtitle = "Atf4") +
  theme(
    plot.subtitle = element_text(face = "italic"),
    legend.position = "none",
    strip.text.x = element_text(face = "bold.italic"),
    strip.background = element_blank(),
    panel.spacing = unit(2, "lines"),
    axis.text.x = element_text(angle = 35, vjust = 1, hjust = 1)
  ) +
  scale_y_continuous(breaks = pretty_breaks(n = 4, min.n = 0))

```

```{r}
marker_gene_umap <- (((hspa5_violin / hspa5_umap + plot_layout(heights = c(0.8, 1)))) |
  (atf4_violin / atf4_umap + plot_layout(heights = c(0.8, 1))) |
  (ddit3_violin / ddit3_umap + plot_layout(heights = c(0.7, 1))))

marker_gene_umap
ggsave( filename = paste(results_dir, "Figure 2EFG.png", sep = ""), width = 9, height = 4.5, device = "png", dpi = 700)
```

## WGCNA

```{r create_metacells}
# WGCNA is performed for variable genes identified by SCTransform
genes.keep <- VariableFeatures(classified_cells, assay = "SCT")

# create metacells for each treatment
seurat_list <- list()
for (group in unique(classified_cells$treatment)) {
  print(group)
  cur_seurat <- subset(classified_cells, treatment == group)
  cur_seurat <- cur_seurat[genes.keep, ]
  cur_metacell_seurat <- scWGCNA::construct_metacells(
    cur_seurat,
    name = group,
    k = 60,
    reduction = "umap",
    assay = "SCT", slot = "data"
  )
  seurat_list[[group]] <- cur_metacell_seurat
}

metacell_seurat <- merge(seurat_list[[1]], seurat_list[2:length(seurat_list)])

# plot showing the number of metacells for each treatment
metacell_counts <- metacell_seurat@meta.data %>%
  ggplot(aes(orig.ident, fill=factor(orig.ident))) +
  geom_bar(width = 0.75) +
  geom_text(aes(label=..count..),
            stat='count',
            size = 3, 
            position = position_stack(vjust = 1.07),
            color="black") +
  labs(x="", y="# Cells", fill="", title="") +
  theme_bw() +
  theme( legend.position = "none") +
  scale_fill_manual(values = viridis_cp)  

# umap for metacell data
#scale data
metacell_plot_data <- ScaleData(metacell_seurat, features = rownames(metacell_seurat))

# PCA
metacell_plot_data  <- RunPCA(metacell_plot_data , features=rownames(metacell_seurat))

# UMAP
metacell_plot_data  <- RunUMAP(metacell_plot_data , reduction='pca', dims=1:15)

# plot
metacell_umap <- DimPlot(metacell_plot_data ,  reduction='umap', label=TRUE)

metacell_umap <- metacell_umap[[1]]$data %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = ident)) +
  geom_point(size = 0.1) +
  theme_bw() +
  coord_fixed() +
  scale_color_manual(values = viridis_cp) +
  labs(x = "UMAP 1", y = "UMAP 2", color = "", subtitle = "") +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  theme(
    plot.title = element_text(size = 10, face = "bold.italic"),
    legend.position = "right",
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

metacell_counts + metacell_umap + plot_annotation(tag_levels = "a") + plot_layout(widths = c(2,1), heights = c(1,1)) 

ggsave(filename = paste(results_dir, "Figure S10.png", sep = ""), width = 9, height = 5, device = "png", dpi = 700)

```

```{r find_soft_tresh}
#nclusters <- length(unique(metacell_seurat$orig.ident))

genes.use <- rownames(metacell_seurat)

targets <- metacell_seurat@meta.data

group <- as.factor(metacell_seurat$orig.ident)

#create expression matrix
datExpr <- as.data.frame(GetAssayData(metacell_seurat, assay = "RNA", slot = "data")[genes.use, ])

# transpose
datExpr <- as.data.frame(t(datExpr))

datExpr <- datExpr[, goodGenes(datExpr)]

# Choose a set of soft-thresholding powers
powers <- c(seq(1, 10, by = 1), seq(12, 50, by = 2))

# assess metrics over selected soft threshold range
powerTable <- list(
  data = pickSoftThreshold(
    datExpr,
    powerVector = powers,
    verbose = 100,
    networkType = "signed",
    corFnc = "cor"
  )[[2]]
)

# plot 
figure_s11a <- powerTable$data %>%
  ggplot(aes(x=Power, y=SFT.R.sq)) +
  geom_point(size=0.1) +
  labs(y="SFT fit", x="Power") +
  geom_vline(xintercept = 7, linetype=11, size=0.2, color="red") +
  theme_bw()
  
figure_s11b <- powerTable$data %>%
  ggplot(aes(x=Power, y=mean.k.)) +
  geom_point(size=0.1) +
  labs(y="Mean connectivity", x="Power") +
  geom_vline(xintercept = 7, linetype=11, size=0.2, color="red") +
  theme_bw() 


figure_s11a + figure_s11b +plot_annotation(tag_levels = "a")

ggsave(filename = paste(results_dir, "Figure S11.png", sep = ""), width = 6, height = 2, device = "png", dpi = 700)
```


```{r run_wgcna}
# selected soft threshold
softPower <- 7

nSets <- 1
setLabels <- "treatment"
shortLabels <- setLabels

multiExpr <- list()
multiExpr[["TM"]] <- list(data = datExpr)

checkSets(multiExpr) # check data size

# construct network
net <- blockwiseConsensusModules(multiExpr,
  blocks = NULL,
  maxBlockSize = 30000, ## This should be set to a smaller size if the user has limited RAM
  randomSeed = 12345,
  corType = "pearson",
  power = softPower,
  consensusQuantile = 0.3,
  networkType = "signed",
  TOMType = "unsigned",
  TOMDenom = "min",
  scaleTOMs = TRUE, scaleQuantile = 0.8,
  sampleForScaling = TRUE, sampleForScalingFactor = 1000,
  useDiskCache = TRUE, chunkSize = NULL,
  deepSplit = 4,
  pamStage = FALSE,
  detectCutHeight = 0.995,
  minModuleSize = 50,
  mergeCutHeight = 0.1,
  saveConsensusTOMs = TRUE,
  consensusTOMFilePattern = "ConsensusTOM-block.%b.rda"
)

consMEs <- net$multiMEs
moduleLabels <- net$colors
table(moduleLabels)

# change the base colors to something better
moduleLabels <- moduleLabels %>%
  revalue(c(
    blue = "#1F78B4",
    brown = "#33A02C",
    magenta = "#E31A1C",
    yellow = "#FFFF99",
    green = "#6A3D9A",
    grey = "#D3D3D3",
    pink = "#FF7F00",
    red = "#B15928",
    turquoise = "#3acabb"
  ))

# Convert the numeric labels to color labels
moduleColors <- as.character(moduleLabels)
length(unique(moduleColors))


consTree <- net$dendrograms[[1]]

# generatate module eigengenes
MEs <- moduleEigengenes(multiExpr[[1]]$data, colors = moduleColors, nPC = 1)$eigengenes
MEs <- orderMEs(MEs)
meInfo <- data.frame(rownames(datExpr), MEs)
colnames(meInfo)[1] <- "SampleID"

# calculate intramodular connectivity
KMEs <- signedKME(datExpr, MEs, outputColumnName = "kME", corFnc = "bicor")

# compile into a module metadata table
geneInfo <- as.data.frame(cbind(colnames(datExpr), moduleColors, KMEs))

# how many modules did we get?
nmodules <- length(unique(moduleColors))

# merged gene symbol column
colnames(geneInfo)[1] <- "GeneSymbol"
colnames(geneInfo)[2] <- "Initially.Assigned.Module.Color"
PCvalues <- MEs


# save dendrogram figure
png(file = paste(results_dir, "Figure_3A.png", sep = ""), units = "in", width = 6, height = 3, res = 700)

plotDendroAndColors(consTree, moduleColors, "Module colors",
  dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05,
  main = paste0("")
)

dev.off()

# show dendrogram figure
plotDendroAndColors(consTree, moduleColors, "Module colors",
  dendroLabels = FALSE, hang = 0.03, addGuide = TRUE, guideHang = 0.05,
  main = paste0("")
)

# figure to show number of genes in each module

# switch the colors back for plot
num_gene_mod_counts <- moduleColors %>%
  revalue(c(
    "#1F78B4" = "blue",
    "#33A02C" = "green",
    "#E31A1C" = "red",
    "#FFFF99" = "yellow",
    "#6A3D9A" = "purple",
    "#D3D3D3" = "grey",
    "#FF7F00" = "orange",
    "#B15928" = "brown",
    "#3acabb" = "turquoise"
  ))

# plot
data.frame(table(num_gene_mod_counts)) %>%
  filter(num_gene_mod_counts != "grey") %>%
  ggplot(aes(x = num_gene_mod_counts, y = Freq, label = Freq, fill = num_gene_mod_counts)) +
  geom_bar(stat = "identity", colour = "grey") +
  geom_label(
    fill = "white",
    size = 3,
    # position = position_stack(vjust = 1.1),
    color = "black"
  ) +
  theme_bw() +
  labs(x = "", y = "# Coexpressed genes") +
  theme(legend.position = "none") +
  scale_fill_manual(values = c(
    "black", "#1F78B4", "#B15928", "#33A02C", "#FF7F00", "#6A3D9A", "#E31A1C",
    "#3acabb", "#FFFF99"
  ))

ggsave(
  filename = paste(results_dir, "Figure 3B.png", sep = ""),
  width = 5,
  height = 3,
  device = "png",
  dpi = 700
)
```


Now we annotate each of the 3,000 genes used for WGCNA analysis.

```{r annotate_wgcna_genes}
ensembl <- useEnsembl(
  biomart = "ensembl",
  dataset = "cgpicr_gene_ensembl",
  version = "102"
)

biomart <- getBM(
  attributes = c(
    "ensembl_gene_id_version", "external_gene_name",
    "description", "gene_biotype", "entrezgene_accession"
  ),
  filters = "ensembl_gene_id_version",
  values = geneInfo$GeneSymbol,
  mart = ensembl,
  uniqueRows = T
)

# new data frame
geneInfo_annotated <- geneInfo

# switch colours back to names
geneInfo_annotated$Initially.Assigned.Module.Color <- revalue(
  geneInfo_annotated$Initially.Assigned.Module.Color,
  c(
    "#1F78B4" = "blue",
    "#33A02C" = "green",
    "#E31A1C" = "red",
    "#FFFF99" = "yellow",
    "#6A3D9A" = "purple",
    "#D3D3D3" = "grey",
    "#FF7F00" = "orange",
    "#B15928" = "brown",
    "#3acabb" = "turquoise"
  )
)

new_colnames_colour_names <- revalue(
  colnames(geneInfo_annotated),
  c(
    "kME#1F78B4" = "kMEblue",
    "kME#33A02C" = "kMEgreen",
    "kME#E31A1C" = "kMEred",
    "kME#FFFF99" = "kMEyellow",
    "kME#6A3D9A" = "kMEpurple",
    "kME#D3D3D3" = "kMEgrey",
    "kME#FF7F00" = "kMEorange",
    "kME#B15928" = "kMEbrown",
    "kME#3acabb" = "kMEturquoise"
  )
)

# make the annotated gene information table
colnames(geneInfo_annotated) <- new_colnames_colour_names

geneInfo_annotated <- geneInfo_annotated %>%
  dplyr::rename(
    ensembl_gene_id_version = GeneSymbol,
    module_color = Initially.Assigned.Module.Color
  ) %>%
  left_join(biomart, by = "ensembl_gene_id_version") %>% # add annotations
  mutate_all(na_if, "") %>%
  distinct(ensembl_gene_id_version, .keep_all = T) %>%
  mutate(gene_symbol = coalesce(external_gene_name, 
                                entrezgene_accession)) %>%
  dplyr::select(c(-entrezgene_accession, 
                  -external_gene_name)) %>%
  dplyr::select(c(ensembl_gene_id_version, 
                  gene_biotype, 
                  gene_symbol, 
                  description, 
                  everything())) %>%
  dplyr::arrange(module_color)





filename <- paste(results_dir, "Table S4.xlsx",
  sep = ""
)

write_xlsx(list(
  a = geneInfo_annotated
),
path = filename,
format_headers = TRUE
)

# show the WGCNA outs
datatable(geneInfo_annotated, rownames = FALSE, filter="top", options = list(pageLength = 5, scrollX=T))
```

Plot the module eigengenes for each module against the sample IDs to assess correlations in expression. 

```{r correlation_with_tm, fig.height=10, fig.width=8}
plot_df <- cbind(dplyr::select(targets, c(orig.ident)), PCvalues)


new_colnames_plot_df <- revalue(colnames(plot_df), 
                           c("ME#1F78B4" = "MEblue",
                             "ME#33A02C" = "MEgreen",
                             "ME#E31A1C" = "MEred",
                             "ME#FFFF99" = "MEyellow",
                             "ME#6A3D9A" = "MEpurple",
                             "ME#D3D3D3" = "MEgrey",
                             "ME#FF7F00" = "MEorange",
                             "ME#B15928" = "MEbrown",
                             "ME#3acabb" = "MEturquoise"))

colnames(plot_df) <- new_colnames_plot_df

plot_df <- reshape2::melt(plot_df, id.vars = c("orig.ident"))

plot_df$orig.ident <- factor(plot_df$orig.ident, 
                             levels = c("TM-", "TM+ 1hr", "TM+ 2hr", "TM+ 4hr", "TM+ 8hr"))

colors <- sub("ME", "", as.character(levels(plot_df$variable)))

wgcna_boxplot_full <- plot_df %>%
  filter(variable != "MEgrey") %>%
  ggplot(aes(x = orig.ident, y = value, fill = orig.ident)) +
  scale_fill_manual(values = c("#fde725", "#5ec962", "#21918c", "#31688e", "#440154")) +
  geom_boxplot(alpha=0.5) +
  RotatedAxis() +
  ylab("Module Eigengene") +
  xlab("") +
  theme_bw() +
  theme(
    legend.position = "none",
    strip.text.x = element_text(face = "bold"),
    strip.background = element_blank(),
    axis.text.x = element_text(angle = 35, vjust = 1, hjust = 1)
  ) +
  facet_wrap(~variable, scales = "free", ncol = 3)

wgcna_boxplot_full
ggsave(plot = wgcna_boxplot_full, filename = paste(results_dir, "Figure S9.png", sep = ""), width = 9, height = 9, device = "png", dpi = 700)
```

The ME expression of the blue, green and red coexpression modules seem to be related to TM treatment over the time course. Let's see if any biological processes are enriched.  

```{r enirchment_analysis}
enrichdir <- paste0(results_dir, "/enrichment_analysis/")
suppressMessages(if (file.exists(enrichdir)) {
  unlink(enrichdir, recursive = T)
})
if (!dir.exists(enrichdir)) {
  dir.create(enrichdir)
}


listGeneSet(organism = "mmusculus")

selected_colors <- c("blue","green","red")

for (color in selected_colors) {
  current_genes <- geneInfo_annotated %>%
    filter(module_color == color) %>%
    dplyr::select(gene_symbol)

  write(current_genes$gene_symbol, file = paste0(enrichdir, color, "_genes.txt"))
}

enrich_result <- WebGestaltRBatch(
  enrichMethod = "ORA",
  organism = "mmusculus",
  enrichDatabase = c("geneontology_Biological_Process"),
  enrichDatabaseType = "genesymbol",
  interestGeneFolder = enrichdir,
  interestGeneType = "genesymbol",
  referenceSet = "genome",
  minNum = 10,
  maxNum = 500,
  sigMethod = "fdr",
  fdrMethod = "BH",
  fdrThr = 0.05,
  topThr = 10,
  reportNum = 20,
  perNum = 1000,
  projectName = "SBO ER stress",
  isOutput = TRUE,
  outputDirectory = enrichdir,
  dagColor = "continuous",
  setCoverNum = 10,
  networkConstructionMethod = NULL,
  neighborNum = 10,
  highlightType = "Seeds",
  highlightSeedNum = 10,
  nThreads = 32
)

filename <- paste(results_dir, "Table S5.xlsx",
  sep = ""
)

write_xlsx(list(
   a=enrich_result[[1]]$enrichResult[1:8, c(1, 2, 4, 5, 7, 8, 9, 11)],
   b=enrich_result[[2]]$enrichResult[1:8, c(1, 2, 4, 5, 7, 8, 9, 11)],
   c=enrich_result[[3]]$enrichResult[1:8, c(1, 2, 4, 5, 7, 8, 9, 11)]
 ),
 path = filename,
 format_headers = TRUE
 )
```


plot the ME and enrichment for blue module genes  
```{r blue_module_plot,fig.height=3, fig.width=10}
wgcna_boxplot_blue <- plot_df %>%
  mutate(new_sample_label=case_when(
    orig.ident == "TM-" ~ "TM-",
    orig.ident == "TM+ 1hr" ~ "TM+\n1hr",
    orig.ident == "TM+ 2hr" ~ "TM+\n2hr",
    orig.ident == "TM+ 4hr" ~ "TM+\n4hr",
    orig.ident == "TM+ 8hr" ~ "TM+\n8hr",
    )) %>%
  filter( variable == "MEblue") %>%
  ggplot(aes(x = new_sample_label, y = value, fill = orig.ident)) +
  scale_fill_manual(values = c("#fde725", "#5ec962", "#21918c", "#31688e", "#440154")) +
  geom_boxplot(alpha=0.5) + 
  labs(x="", y="Module Eigengene", title = "Blue Module") +
  RotatedAxis() +
  theme_bw() +
  theme(
    legend.position = "none"
  ) 

blue_module_enrichment_plot <- enrich_result[[1]]$enrichResult[1:8, c(1, 2, 4, 5, 7, 8, 9, 11)] %>%
  mutate(type = "MEblue") %>%
mutate(FDR = case_when(FDR == 0 ~ 2.2e-16, TRUE ~ FDR)) %>%
  mutate(new_description = str_wrap(description, width = 50)) %>%
  mutate(new_order = fct_reorder(new_description, FDR)) %>%
  ggplot(aes(y = -log10(FDR), x =new_order, fill=type, label=overlap)) +
  geom_bar(stat="identity", width = 0.75) +
     scale_y_continuous(expand = c(0, 0), limits = c(0, -log10(2.2e-16)+ 0.05)) +
  geom_text(aes(label = overlap), vjust = 0.5,hjust = 1.1, colour = "white") +
  #scale_fill_viridis() +
  theme_bw() +
  labs(x = "",y = "-log10(FDR)", title = "GO Biological Process") +
  scale_fill_manual(values = "#1F78B4") +
  theme(strip.text.x = element_text(face = "bold"),strip.background = element_blank(), legend.position = "none",
        axis.text.x = element_text( size = 6),
        axis.text.y = element_text( size = 9),
        panel.spacing = unit(2, "lines")) +
    scale_x_discrete(limits=rev) +
  coord_flip()

fig3_cd <- wgcna_boxplot_blue + blue_module_enrichment_plot + plot_layout(widths=c(1,2.5))
fig3_cd
```

```{r green_module_plot, fig.height=3, fig.width=10}
wgcna_boxplot_green <- plot_df %>%
  mutate(new_sample_label=case_when(
    orig.ident == "TM-" ~ "TM-",
    orig.ident == "TM+ 1hr" ~ "TM+\n1hr",
    orig.ident == "TM+ 2hr" ~ "TM+\n2hr",
    orig.ident == "TM+ 4hr" ~ "TM+\n4hr",
    orig.ident == "TM+ 8hr" ~ "TM+\n8hr",
    )) %>%
  filter( variable == "MEgreen") %>%
  ggplot(aes(x = new_sample_label, y = value, fill = orig.ident)) +
  scale_fill_manual(values = c("#fde725", "#5ec962", "#21918c", "#31688e", "#440154")) +
  geom_boxplot(alpha=0.5) + 
  labs(x="", y="Module Eigengene", title = "Green Module") +
  RotatedAxis() +
  theme_bw() +
  theme(
    legend.position = "none"
  ) 

green_module_enrichment_plot <- enrich_result[[2]]$enrichResult[1:8, c(1, 2, 4, 5, 7, 8, 9, 11)] %>%
  mutate(type = "MEgreen") %>%
mutate(FDR = case_when(FDR == 0 ~ 2.2e-16, TRUE ~ FDR)) %>%
  mutate(new_description = str_wrap(description, width = 50)) %>%
  mutate(new_order = fct_reorder(new_description, FDR)) %>%
  ggplot(aes(y = -log10(FDR), x =new_order, fill=type, label=overlap)) +
  geom_bar(stat="identity", width = 0.75) +
     scale_y_continuous(expand = c(0, 0), limits = c(0, -log10(1.012793e-08)+0.05)) +
  geom_text(aes(label = overlap), vjust = 0.5,hjust = 1.1, colour = "white") +
  #scale_fill_viridis() +
  theme_bw() +
  labs(x = "",y = "-log10(FDR)", title = "GO Biological Process") +
  scale_fill_manual(values = "#33A02C") +
  theme(strip.text.x = element_text(face = "bold"),strip.background = element_blank(), legend.position = "none",
        axis.text.x = element_text( size = 6),
        panel.spacing = unit(2, "lines")) +
    scale_x_discrete(limits=rev) +
  coord_flip()

fig3_ef <- wgcna_boxplot_green + green_module_enrichment_plot + plot_layout(widths=c(1,2.5))
fig3_ef
```

```{r red_module_plot, fig.height=3, fig.width=10}
wgcna_boxplot_red <- plot_df %>%
  mutate(new_sample_label=case_when(
    orig.ident == "TM-" ~ "TM-",
    orig.ident == "TM+ 1hr" ~ "TM+\n1hr",
    orig.ident == "TM+ 2hr" ~ "TM+\n2hr",
    orig.ident == "TM+ 4hr" ~ "TM+\n4hr",
    orig.ident == "TM+ 8hr" ~ "TM+\n8hr",
    )) %>%
  filter( variable == "MEred") %>%
  ggplot(aes(x = new_sample_label, y = value, fill = orig.ident)) +
  scale_fill_manual(values = c("#fde725", "#5ec962", "#21918c", "#31688e", "#440154")) +
  geom_boxplot(alpha=0.5) + 
  labs(x="", y="Module Eigengene", title = "Red Module") +
  RotatedAxis() +
  theme_bw() +
  theme(
    legend.position = "none"
  ) 

red_module_enrichment_plot <- enrich_result[[3]]$enrichResult[1:8, c(1, 2, 4, 5, 7, 8, 9, 11)] %>%
  mutate(type = "MEred") %>%
mutate(FDR = case_when(FDR == 0 ~ 2.2e-16, TRUE ~ FDR)) %>%
  mutate(new_description = str_wrap(description, width = 50)) %>%
  mutate(new_order = fct_reorder(new_description, FDR)) %>%
  ggplot(aes(y = -log10(FDR), x =new_order, fill=type, label=overlap)) +
  geom_bar(stat="identity", width = 0.75) +
     scale_y_continuous(expand = c(0, 0), limits = c(0, -log10(3.455569e-10)+0.05)) +
  geom_text(aes(label = overlap), vjust = 0.5,hjust = 1.1, colour = "white") +
  #scale_fill_viridis() +
  theme_bw() +
  labs(x = "",y = "-log10(FDR)", title = "GO Biological Process") +
  scale_fill_manual(values = "#E31A1C") +
  theme(strip.text.x = element_text(face = "bold"),strip.background = element_blank(), legend.position = "none",
        axis.text.x = element_text( size = 6),
        panel.spacing = unit(2, "lines")) +
    scale_x_discrete(limits=rev) +
  coord_flip()

fig3_gh <- wgcna_boxplot_red + red_module_enrichment_plot + plot_layout(widths=c(1,2.5))
fig3_gh
```

```{r manuscript_figure, fig.height=9.75, fig.width=10}
fig3_cd / fig3_ef / fig3_gh

ggsave(filename = paste(results_dir, "Figure 3CDEFGH.png", sep = ""), width = 10, height = 9.75, device = "png", dpi = 700)
```

```{r blue_module_umaps, fig.height=7, fig.width=9}
selected_genes <-geneInfo_annotated %>%
 # filter(gene_symbol %in% cell_cycle_phase_transition) %>%
  arrange(-kMEblue) %>%
  top_n(9, kMEblue)
  
for (i in 1:dim(selected_genes)[1]) {

   umap_data <- FeaturePlot(classified_cells, features = selected_genes$ensembl_gene_id_version[i])$data %>%
  mutate(gene = selected_genes$gene_symbol[i]) %>%
   dplyr::rename(expression = selected_genes$ensembl_gene_id_version[i])
 
umap_plot <- umap_data %>%
  filter(gene == selected_genes$gene_symbol[i]) %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = expression)) +
  geom_point(size = 0.1) +
  scale_color_viridis(option = "turbo") +
  theme_bw() +
  #scale_x_reverse() +
  labs(x = "UMAP 2", y = "UMAP 1", color = paste0(selected_genes$gene_symbol[i],"\nExpression \nlevel"), 
       title= selected_genes$gene_symbol[i], 
       subtitle = paste0(" kMEblue = ", signif(selected_genes$kMEblue[i],digits = 3))) +
  theme(
    plot.title = element_text(face = "italic", size=12),
    plot.subtitle = element_text(size=10),
    legend.title = element_text(size=10)
    ) 

assign(paste0("fig_s13", LETTERS[i]), umap_plot)
}

(fig_s13A + fig_s13B + fig_s13C) / 
  (fig_s13D + fig_s13E + fig_s13F) /
    (fig_s13G + fig_s13H + fig_s13I) 

ggsave(filename = paste(results_dir, "Figure S13.png", sep = ""), width = 9, height = 7, device = "png", dpi = 700)
```

```{r green_module_umaps, fig.height=7, fig.width=9}
selected_genes <-geneInfo_annotated %>%
 # filter(gene_symbol %in% cell_cycle_phase_transition) %>%
  arrange(-kMEgreen) %>%
  top_n(9, kMEgreen)
  
for (i in 1:dim(selected_genes)[1]) {

   umap_data <- FeaturePlot(classified_cells, features = selected_genes$ensembl_gene_id_version[i])$data %>%
  mutate(gene = selected_genes$gene_symbol[i]) %>%
   dplyr::rename(expression = selected_genes$ensembl_gene_id_version[i])
 
umap_plot <- umap_data %>%
  filter(gene == selected_genes$gene_symbol[i]) %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = expression)) +
  geom_point(size = 0.1) +
  scale_color_viridis(option = "turbo") +
  theme_bw() +
  #scale_x_reverse() +
  labs(x = "UMAP 2", y = "UMAP 1", color = paste0(selected_genes$gene_symbol[i],"\nExpression \nlevel"), 
       title= selected_genes$gene_symbol[i], 
       subtitle = paste0(" kMEgreen = ", signif(selected_genes$kMEgreen[i],digits = 3))) +
  theme(
    plot.title = element_text(face = "italic", size=12),
    plot.subtitle = element_text(size=10),
    legend.title = element_text(size=10)
    ) 

assign(paste0("fig_s14", LETTERS[i]), umap_plot)
}

(fig_s14A + fig_s14B + fig_s14C) / 
  (fig_s14D + fig_s14E + fig_s14F) /
    (fig_s14G + fig_s14H + fig_s14I) 

ggsave(filename = paste(results_dir, "Figure S14.png", sep = ""), width = 9, height = 7, device = "png", dpi = 700)
```

```{r red_module_umaps, fig.height=7, fig.width=9}
selected_genes <-geneInfo_annotated %>%
  #filter(gene_symbol %in% c("Hmgcs1","Hmgcr","Mvd","Pmvk","Fdft1","Msmo1", "Erg28", "Slc27a3", "Gamt")) %>%
  filter(!is.na(gene_symbol)) %>%
  arrange(-kMEred) %>%
  top_n(9, kMEred)
  
for (i in 1:dim(selected_genes)[1]) {

   umap_data <- FeaturePlot(classified_cells, features = selected_genes$ensembl_gene_id_version[i])$data %>%
  mutate(gene = selected_genes$gene_symbol[i]) %>%
   dplyr::rename(expression = selected_genes$ensembl_gene_id_version[i])
 
umap_plot <- umap_data %>%
  filter(gene == selected_genes$gene_symbol[i]) %>%
  ggplot(aes(x = UMAP_1, y = UMAP_2, color = expression)) +
  geom_point(size = 0.1) +
  scale_color_viridis(option = "turbo") +
  theme_bw() +
  #scale_x_reverse() +
  labs(x = "UMAP 2", y = "UMAP 1", color = paste0(selected_genes$gene_symbol[i],"\nExpression \nlevel"), 
       title= selected_genes$gene_symbol[i], 
       subtitle = paste0(" kMEred = ", signif(selected_genes$kMEred[i],digits = 3))) +
  theme(
    plot.title = element_text(face = "italic", size=12),
    plot.subtitle = element_text(size=10),
    legend.title = element_text(size=10)
    ) 

assign(paste0("fig_s15", LETTERS[i]), umap_plot)
}

(fig_s15A + fig_s15B + fig_s15C) / 
  (fig_s15D + fig_s15E + fig_s15F) /
    (fig_s15G + fig_s15H + fig_s15I) 

ggsave(filename = paste(results_dir, "Figure S15.png", sep = ""), width = 9, height = 7, device = "png", dpi = 700)
```

```{r}
geneInfo_annotated %>%
  filter(module_color == "red") %>%
  arrange(gene_symbol)
```


```{r show_session_info}
sessionInfo()
```
